Identifying `rtrees` Training Data

### 2020-09-29 journals/09-identifying_rtrees_training_rdata/09-identifying_rtrees_training_data.Rmd
renamed files to match other journals

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE, 
  dpi = 300
)
```

@@

```{r}
length(unique(c(CCFs_withlands$land_id1, CCFs_withlands$land_id2)))
```

@@

@@

```{r}
rtrees_pred = predict(bulletxtrctr::rtrees, type = "prob")[,2]
hc_pred = predict(bulletxtrctr::rtrees, hamby_comparisons %>% select(rtrees_features), type = "prob")[,2]
ccfwl_pred = predict(bulletxtrctr::rtrees, CCFs_withlands %>% select(rtrees_features), type = "prob")[,2]

@@

```{r fig.width = 12, fig.height = 5}
plot_grid(
  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees2, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Retrained RF on Hamby comparison data versus rtrees; Seed 20140501"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees3, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees3 votes",
         title = "Retrained RF on Hamby comparison data versus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees2, y = rtrees3, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Both retrained RFs on Hamby comparison data"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees4, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees4 votes",
         title = "Retrained RF on CCFs with lands data versus rtrees; Seed 20140501"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees5, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Retrained RF on CCFs with lands data versus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees4, y = rtrees5, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Both retrained RFs on CCFs with lands data"),
  nrow = 2
)
```

@@

```{r fig.width = 12, fig.height = 5}
plot_grid(
  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees2), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Retrained RF on Hamby comparison data versus rtrees; Seed 20140501"),

  rtrees_votes%>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees3), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees3 votes",
         title = "Retrained RF on Hamby comparison data versus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees2), y = sort(rtrees3), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Both retrained RFs on Hamby comparison data"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees4), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees4 votes",
         title = "Retrained RF on CCFs with lands data versus rtrees; Seed 20140501"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees5), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Retrained RF on CCFs with lands data versus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees4), y = sort(rtrees5), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Both retrained RFs on CCFs with lands data"),
  nrow = 2
)
```

@@

```{r}
hamby_comparison_importance <-
  map_df(.x = rfs_hamby_comparison, .f = extract_importance, .id = "rf_model_id")
CCFs_withlands_importance <-
  map_df(.x = rfs_CCFs_withlands, .f = extract_importance, .id = "rf_model_id")
```

@@

```{r}
rtrees_importance <- 
  data.frame(features = rownames(bulletxtrctr::rtrees$importance),
             bulletxtrctr::rtrees$importance) %>%
  arrange(desc(MeanDecreaseGini)) %>%
  mutate(rank = 1:n(),
         rf_model_id = "rtrees") %>%
  rename("importance" = "MeanDecreaseGini") %>%
  select(rf_model_id, features, importance, rank)
```

@@

Histograms of the class errors (true and false) show that the rtrees values appear to be in the extreme (false) or outside of the new random forest models (true):

```{r fig.width = 10, fig.height = 4.5}
hist_true_hc <-
  hamby_comparison_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false_hc <- 
  hamby_comparison_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_false, color = model))

hist_true_ccfwl <-
  CCFs_withlands_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false_ccfwl <- 
  CCFs_withlands_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_false, color = model))

@@

Heatmap depicting feature importance across models:

```{r fig.width = 12, fig.height = 5}
plot_grid(
  bind_rows(rtrees_importance, hamby_comparison_importance) %>%
    mutate(
      rf_model_id = factor(rf_model_id, levels = c("rtrees", 1:length(seeds))),
      rank = factor(rank, levels = 1:9),
      features = factor(features, levels = rev(rtrees_importance$features))
    ) %>%
    ggplot(aes(x = rf_model_id, y = features, fill = rank)) +
    geom_tile() +
    scale_fill_brewer(palette = "RdYlBu") + 
    theme_bw() +
    labs(
      x = "Random Forest Model", 
      y = "Feature", 
      title = "Comparing rtrees importance to random forests fit using Hamby comparisons data"
    ),

  bind_rows(rtrees_importance, CCFs_withlands_importance) %>%
    mutate(
      rf_model_id = factor(rf_model_id, levels = c("rtrees", 1:length(seeds))),
      rank = factor(rank, levels = 1:9),
      features = factor(features, levels = rev(rtrees_importance$features))
    ) %>%
    ggplot(aes(x = rf_model_id, y = features, fill = rank)) +
    geom_tile() +
    scale_fill_brewer(palette = "RdYlBu") + 
    theme_bw() +
    labs(
      x = "Random Forest Model", 
      y = "Feature", 
      title = "Comparing rtrees importance to random forests fit using CCFs with lands data"
    )
)
```

```{r fig.width = 12, fig.height = 5}
plot_grid(
  bind_rows(rtrees_importance, hamby_comparison_importance) %>%
    mutate(
      rf_model_id = factor(rf_model_id, levels = c("rtrees", 1:length(seeds))),
      rank = factor(rank, levels = 1:9),
      features = factor(features, levels = rev(rtrees_importance$features)),
      rtrees_ind = ifelse(rf_model_id == "rtrees", "rtrees", "Hamby Comparison")
    ) %>%
    ggplot(aes(x = rank, y = features, group = rf_model_id, color = rtrees_ind)) +
    geom_point(alpha = 0.5, shape = 1) + 
    geom_line(alpha = 0.5) +
    scale_color_manual(values = c("grey50", "blue")) +
    theme_bw() +
    labs(
      x = "Random Forest Model", 
      y = "Feature", 
      title = "Comparing rtrees importance to random forests fit using Hamby comparisons data"
    ), 
    bind_rows(rtrees_importance, CCFs_withlands_importance) %>%
    mutate(
      rf_model_id = factor(rf_model_id, levels = c("rtrees", 1:length(seeds))),
      rank = factor(rank, levels = 1:9),
      features = factor(features, levels = rev(rtrees_importance$features)),
      rtrees_ind = ifelse(rf_model_id == "rtrees", "rtrees", "CCFs with lands")
    ) %>%
    ggplot(aes(x = rank, y = features, group = rf_model_id, color = rtrees_ind)) +
    geom_point(alpha = 0.5, shape = 1) + 
    geom_line(alpha = 0.5) +
    scale_color_manual(values = c("grey50", "blue")) +
    theme_bw() +
    labs(
      x = "Random Forest Model", 
      y = "Feature", 
      title = "Comparing rtrees importance to random forests fit using CCFs with lands data"
    )
)
```

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE, 
  dpi = 500
)
```

@@

Save the cleaned data:

```{r}
write.csv(x = hamby_comparisons, file = "../../../data/hamby_comparisons_clean.csv", row.names = FALSE)
write.csv(x = CCFs_withlands, file = "../../../data/CCFs_withlands_clean.csv", row.names = FALSE)
```

@@

Save the cleaned and filtered data:

```{r}
write.csv(x = hamby_comparisons_filtered , file = "../../../data/hamby_comparisons_filtered.csv", row.names = FALSE)
```

@@

```{r}
rtrees_pred = predict(bulletxtrctr::rtrees, type = "prob")[,2]
hc_pred = predict(bulletxtrctr::rtrees, hamby_comparisons %>% select(all_of(rtrees_features)), type = "prob")[,2]
ccfwl_pred = predict(bulletxtrctr::rtrees, CCFs_withlands %>% select(all_of(rtrees_features)), type = "prob")[,2]

@@

```{r fig.width = 15, fig.height = 8}
plot_grid(
  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees2, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Retrained RF on Hamby comparison data \nversus rtrees; Seed 20140501"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees3, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees3 votes",
         title = "Retrained RF on Hamby comparison data \nversus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees2, y = rtrees3, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Both retrained RFs on Hamby \ncomparison data"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees4, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees4 votes",
         title = "Retrained RF on CCFs with lands data \nversus rtrees; Seed 20140501"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees, y = rtrees5, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Retrained RF on CCFs with lands data \nversus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = rtrees4, y = rtrees5, color = label
    )) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Both retrained RFs on CCFs with \nlands data"),
  nrow = 2
)
```

@@

```{r fig.width = 15, fig.height = 8}
plot_grid(
  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees2), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Retrained RF on Hamby comparison data \nversus rtrees; Seed 20140501"),

  rtrees_votes%>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees3), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees3 votes",
         title = "Retrained RF on Hamby comparison data \nversus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees2), y = sort(rtrees3), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees2 votes",
         title = "Both retrained RFs on Hamby \ncomparison data"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees4), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees4 votes",
         title = "Retrained RF on CCFs with lands data \nversus rtrees; Seed 20140501"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees), y = sort(rtrees5), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Retrained RF on CCFs with lands data \nversus rtrees; Seed 20170222"),

  rtrees_votes %>%
    ggplot(aes(
      x = sort(rtrees4), y = sort(rtrees5), color = label
    )) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 1) +
    labs(x = "rtrees votes",
         y = "rtrees5 votes",
         title = "Both retrained RFs on CCFs with \nlands data"),
  nrow = 2
)
```

@@

```{r}
hamby_comparison_importance <-
  map_df(.x = rfs_hamby_comparison, .f = extract_importance, .id = "rf_model_id") %>%
  mutate(scenario = "Hamby Comparisons")
CCFs_withlands_importance <-
  map_df(.x = rfs_CCFs_withlands, .f = extract_importance, .id = "rf_model_id") %>%
  mutate(scenario = "CCFs with lands")
```

@@

```{r}
rtrees_importance <- 
  data.frame(features = rownames(bulletxtrctr::rtrees$importance),
             bulletxtrctr::rtrees$importance) %>%
  arrange(desc(MeanDecreaseGini)) %>%
  mutate(rank = 1:n(),
         rf_model_id = "rtrees", 
         scenario = "rtrees") %>%
  rename("importance" = "MeanDecreaseGini") %>%
  select(rf_model_id, features, importance, rank, scenario)
```

@@

Histograms of the class errors (true and false) show that the rtrees class errors are in the tails of the distributions of the class errors from the models trained on the `hamby_comparisons` data  but closer to the center of the class error distributions from the models trained on the `CCFs_withlands` data:

```{r fig.width = 10, fig.height = 4.5}
hist_true_hc <-
  hamby_comparison_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram(bins = 20) +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false_hc <- 
  hamby_comparison_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram(bins = 20) +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_false, color = model))

hist_true_ccfwl <-
  CCFs_withlands_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram(bins = 20) +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false_ccfwl <- 
  CCFs_withlands_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram(bins = 20) +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_false, color = model))

@@

Heatmap depicting feature importance across models (note that `rtrees` importance ranks for `D` and `non_cms` are more similar to random forests trained using `CCFs_withlands` than `hamby_comparisons`):

```{r fig.width = 15, fig.height = 5}
bind_rows(rtrees_importance, hamby_comparison_importance, CCFs_withlands_importance) %>%
    mutate(
      scenario = factor(scenario, levels = c("rtrees", "Hamby Comparisons", "CCFs with lands")),
      rf_model_id = factor(rf_model_id, levels = c("rtrees", 1:length(seeds))),
      rank = factor(rank, levels = 1:9),
      features = factor(features, levels = rev(rtrees_importance$features))
    ) %>%
    ggplot(aes(x = rf_model_id, y = features, fill = rank)) +
    geom_tile() +
    facet_grid(. ~ scenario, scales = "free_x") +
    scale_fill_brewer(palette = "RdYlBu") + 
    theme_bw() +
    labs(
      x = "Random Forest Model", 
      y = "Feature", 
      title = "Comparing Feature Importance to rtrees"
    )
```

Plots of features versus importance ranks (importance ranks vary more in the middle of the features for the models trained using `hamby_comparisons` and near the tail end of the features for the models trained using `CCFs_withlands`):

```{r fig.width = 15, fig.height = 5}
plot_grid(
  bind_rows(rtrees_importance, hamby_comparison_importance) %>%
    mutate(
      rf_model_id = factor(rf_model_id, levels = c("rtrees", 1:length(seeds))),
      rank = factor(rank, levels = 1:9),
      features = factor(features, levels = rev(rtrees_importance$features)),
      rtrees_ind = ifelse(rf_model_id == "rtrees", "rtrees", "Hamby Comparison")
    ) %>%
    arrange(desc(rf_model_id)) %>%
    ggplot(
      aes(
        x = rank,
        y = features,
        group = rf_model_id,
        color = rtrees_ind,
        size = rtrees_ind,
        alpha = rtrees_ind
      )
    ) +
    geom_line() +
    scale_color_manual(values = c("black", "purple")) +
    theme_bw() +
    labs(
      x = "Random Forest Model", 
      y = "Feature", 
      title = "Comparing rtrees importance to random forests fit \nusing Hamby comparisons data"
    ), 
    bind_rows(rtrees_importance, CCFs_withlands_importance) %>%
    mutate(
      rf_model_id = factor(rf_model_id, levels = c("rtrees", 1:length(seeds))),
      rank = factor(rank, levels = 1:9),
      features = factor(features, levels = rev(rtrees_importance$features)),
      rtrees_ind = ifelse(rf_model_id == "rtrees", "rtrees", "CCFs with lands")
    ) %>%
    arrange(desc(rf_model_id)) %>%
    ggplot(
      aes(
        x = rank,
        y = features,
        group = rf_model_id,
        color = rtrees_ind,
        size = rtrees_ind,
        alpha = rtrees_ind
      )
    ) +
    geom_line() +
    scale_color_manual(values = c("black", "purple")) +
    theme_bw() +
    labs(x = "Random Forest Model",
         y = "Feature",
         title = "Comparing rtrees importance to random forests fit \nusing CCFs with lands data")
)
```
